---
title: "Authentication"
---

# Authentication

AutoDesk uses OAuth-based authentication for access to their services. To get started, visit the **Create an App** tutorial at [https://aps.autodesk.com/en/docs/oauth/v2/tutorials/create-app/](https://aps.autodesk.com/en/docs/oauth/v2/tutorials/create-app/) for instructions on creating an app and getting a Client ID and Secret.

## Store Credentials Safely

Store your Client ID, Secret, and access tokens in `~/.Renviron` and access them with `Sys.getenv()` to keep credentials out of your code and version control.

Create or edit `~/.Renviron` with the following entries:

```
client_id = "YOUR_CLIENT_ID"
client_secret = "YOUR_CLIENT_SECRET"
token = "YOUR_ACCESS_TOKEN"
urn = "YOUR_URN"
```

Reload the file without restarting R:

```{r}
#| eval: false
readRenviron("~/.Renviron")
```

For more on this pattern see [Wrapping APIs](https://httr2.r-lib.org/articles/wrapping-apis.html) in the httr2 documentation.

## Get an Access Token

Use `getToken()` to obtain a 2-legged OAuth token. The `scope` parameter controls which APIs the token can access — request only the scopes your task requires.

```{r}
#| eval: false
resp    <- getToken(id = Sys.getenv("client_id"), secret = Sys.getenv("client_secret"))
myToken <- resp$content$access_token
myToken
#> [1] "eyJhbGciOiJSUzI1NiIsImtpZCI6IlU3c1BKNVVpOWNaVjlIR2FhX1pIeDhEd3VYSHcifQ.
#>       eyJjbGllbnRfaWQiOiJBQkNERUYxMjM0NTYiLCJleHAiOjE3NDU0MDA0MDAsInNjb3BlIjpb
#>       ImRhdGE6cmVhZCJdfQ.SflKxwRJSMeKKF2QT4fwpMeJf36POk6yJV_adQssw5c"
```

The full response object also contains the token type and expiry:

```{r}
#| eval: false
str(resp$content)
#> List of 3
#>  $ access_token: chr "eyJhbGciOiJSUzI1NiIsImtpZCI6IlU3c1BKNVVpOWNaVjlIR..."
#>  $ token_type  : chr "Bearer"
#>  $ expires_in  : int 3600
```

Tokens expire after **3600 seconds** (1 hour). Re-call `getToken()` to refresh.

::: callout-warning
**Tokens expire after 1 hour.** If you receive a `401 Unauthorized` error mid-session, your token has expired. Re-call `getToken()` and replace `myToken`. A 401 response looks like:

```{r}
#| eval: false
resp$status_code
#> [1] 401
resp$content$errorCode
#> [1] "AUTH-001"
resp$content$developerMessage
#> [1] "The provided access token has expired."
```
:::

### Available Scopes

| Scope | Description |
|---|---|
| `data:read` | Read files and derivatives |
| `data:write` | Upload and modify files |
| `data:create` | Create new files |
| `data:search` | Search within files |
| `bucket:create` | Create storage buckets |
| `bucket:read` | View bucket details |
| `bucket:update` | Modify bucket settings |
| `bucket:delete` | Delete buckets |
| `code:all` | Access Design Automation APIs |
| `account:read` | Read account information |
| `account:write` | Modify account settings |
| `user-profile:read` | Read user profile data |

Pass multiple scopes as a space-separated string:

```{r}
#| eval: false
resp <- getToken(
  id     = Sys.getenv("client_id"),
  secret = Sys.getenv("client_secret"),
  scope  = "bucket:create bucket:read data:write"
)
myToken <- resp$content$access_token
myToken
#> [1] "eyJhbGciOiJSUzI1NiIsImtpZCI6IlU3c1BKNVVpOWNaVjlIR2FhX1pIeDhEd3VYSHcifQ.
#>       eyJjbGllbnRfaWQiOiJBQkNERUYxMjM0NTYiLCJleHAiOjE3NDU0MDA0MDAsInNjb3BlIjpb
#>       ImJ1Y2tldDpjcmVhdGUiLCJidWNrZXQ6cmVhZCIsImRhdGE6d3JpdGUiXX0.Tq3vK2mP9r"
```

## Rate Limits

APS enforces per-application rate limits that vary by endpoint (typically 60–300 requests per minute). When a limit is exceeded the API returns **HTTP 429**:

```{r}
#| eval: false
resp$status_code
#> [1] 429
resp$content$errorCode
#> [1] "TOO-MANY-REQUESTS"
resp$content$developerMessage
#> [1] "Rate limit exceeded. Retry after 60 seconds."
```

Use the retry helper below around any call that may be polled frequently (e.g. `checkFile()`, `checkPdf()`):

```{r}
#| eval: false
aps_retry <- function(fn, max_tries = 3, wait = 10) {
  for (i in seq_len(max_tries)) {
    resp <- fn()
    if (!identical(resp$status_code, 429L)) return(resp)
    message("Rate limited — waiting ", wait, "s (attempt ", i, "/", max_tries, ")")
    Sys.sleep(wait)
  }
  stop("Max retries exceeded after ", max_tries, " attempts.")
}

# Example: poll translation status with automatic backoff
resp <- aps_retry(function() checkFile(urn = myEncodedUrn, token = myToken))
```
