---
title: "Data Management"
---

# Data Management

The Data Management API provides a way to store and access design files across the AutoDesk Platform Services (APS).

## Create a Bucket

Buckets are named containers for your files. First get a token with the appropriate scopes, then call `makeBucket()`.

```{r}
#| eval: false
resp <- getToken(
  id     = Sys.getenv("client_id"),
  secret = Sys.getenv("client_secret"),
  scope  = "bucket:create bucket:read data:write"
)
myToken <- resp$content$access_token
#> [1] "eyJhbGciOiJSUzI1NiIsImtpZCI6IlU3c1BKNVVpOWNaVjlIR2FhX1pIeDhEd3VYSHcifQ..."
```

Use `makeBucket()` to create a bucket. The `policy` parameter controls how long objects are retained:

| Policy | Retention |
|---|---|
| `"transient"` | 24 hours (default) |
| `"temporary"` | 30 days |
| `"persistent"` | Until explicitly deleted |

::: callout-warning
**Bucket names are globally unique** across every APS application. A name like `"test"` or `"mybucket"` is almost certainly taken. Use a prefix that ties the name to your app or organisation, e.g. `"mycompany-myapp-prod"`. Attempting to create a duplicate returns a `409 Conflict`:

```{r}
#| eval: false
resp$status_code
#> [1] 409
resp$content$reason
#> [1] "Bucket already exists"
```
:::

```{r}
#| eval: false
resp <- makeBucket(token = myToken, bucket = "mybucket", policy = "persistent")
resp$content
#> $bucketKey
#> [1] "mybucket"
#>
#> $bucketOwner
#> [1] "3MVG99OxTyEMCQ3gNp2PjtXbmT8T_MnfCTtpQkuiJNxGtXuoKP7"
#>
#> $createdDate
#> [1] 1745356800000
#>
#> $permissions
#> $permissions[[1]]$authId
#> [1] "3MVG99OxTyEMCQ3gNp2PjtXbmT8T_MnfCTtpQkuiJNxGtXuoKP7"
#>
#> $permissions[[1]]$access
#> [1] "full"
#>
#> $policyKey
#> [1] "persistent"
```

To check the status of a bucket:

```{r}
#| eval: false
resp <- checkBucket(token = myToken, bucket = "mybucket")
resp$content
#> $bucketKey
#> [1] "mybucket"
#>
#> $bucketOwner
#> [1] "3MVG99OxTyEMCQ3gNp2PjtXbmT8T_MnfCTtpQkuiJNxGtXuoKP7"
#>
#> $createdDate
#> [1] 1745356800000
#>
#> $policyKey
#> [1] "persistent"
```

## Upload a File

Use `uploadFile()` for files up to 100 MB. It returns an object containing the `bucketKey`, `objectId` (the URN), `objectKey` (filename), `size`, `contentType`, and `location`. Note the URN and store it in `.Renviron` for future use.

```{r}
#| eval: false
resp <- uploadFile(
  file   = system.file("samples/aerial.dwg", package = "AutoDeskR"),
  token  = myToken,
  bucket = "mybucket"
)
myUrn <- resp$content$objectId
resp$content
#> $bucketKey
#> [1] "mybucket"
#>
#> $objectId
#> [1] "urn:adsk.objects:os.object:mybucket/aerial.dwg"
#>
#> $objectKey
#> [1] "aerial.dwg"
#>
#> $sha1
#> [1] "7f3d9b2e1a6c4f8e0d5b3a9c7e2f1d4b8a6c0e3"
#>
#> $size
#> [1] 3145728
#>
#> $contentType
#> [1] "application/octet-stream"
#>
#> $location
#> [1] "https://developer.api.autodesk.com/oss/v2/buckets/mybucket/objects/aerial.dwg"
```

### Upload Large Files (>100 MB)

For files over 100 MB, use `uploadFileSigned()`, which transfers the file via signed S3 URLs to avoid timeout limits.

```{r}
#| eval: false
resp <- uploadFileSigned(
  file   = "path/to/large_model.rvt",
  token  = myToken,
  bucket = "mybucket"
)
myUrn <- resp$content$objectId
resp$content
#> $bucketKey
#> [1] "mybucket"
#>
#> $objectId
#> [1] "urn:adsk.objects:os.object:mybucket/large_model.rvt"
#>
#> $objectKey
#> [1] "large_model.rvt"
#>
#> $size
#> [1] 524288000
#>
#> $contentType
#> [1] "application/octet-stream"
#>
#> $location
#> [1] "https://developer.api.autodesk.com/oss/v2/buckets/mybucket/objects/large_model.rvt"
```

## List Buckets and Objects

List all buckets associated with your app. Use `limit` to control page size and `startAt` to paginate through results:

```{r}
#| eval: false
resp <- listBuckets(token = myToken, limit = 10)
resp$content$items
#>                          bucketKey                                    createdDate  policyKey
#> 1                       mybucket                                  1745356800000 persistent
#> 2                  archive-2025                                  1740000000000  temporary
#> 3              visualization-dev                                  1738000000000  transient

# Paginate to the next page
resp <- listBuckets(token = myToken, limit = 10, startAt = "visualization-dev")
resp$content$items
#>              bucketKey      createdDate policyKey
#> 1        test-sandbox    1735000000000 transient
```

Use `region` to filter buckets by geography (`"US"` or `"EMEA"`):

```{r}
#| eval: false
resp <- listBuckets(token = myToken, region = "EMEA")
resp$content$items
#>       bucketKey      createdDate  policyKey
#> 1  emea-models    1742000000000 persistent
#> 2  emea-backup    1740000000000  temporary
```

List objects stored in a specific bucket:

```{r}
#| eval: false
resp <- listObjects(token = myToken, bucket = "mybucket", limit = 10)
resp$content$items
#>          objectKey                                              objectId   size
#> 1       aerial.dwg  urn:adsk.objects:os.object:mybucket/aerial.dwg      3145728
#> 2  office_plan.rvt  urn:adsk.objects:os.object:mybucket/office_plan.rvt 52428800
#> 3      bridge.ifc   urn:adsk.objects:os.object:mybucket/bridge.ifc      8388608
```

## Delete Objects and Buckets

Delete a specific object from a bucket:

```{r}
#| eval: false
resp <- deleteObject(token = myToken, bucket = "mybucket", object = "aerial.dwg")
resp$status_code
#> [1] 200
```

Delete an entire bucket (the bucket must be empty first):

```{r}
#| eval: false
resp <- deleteBucket(token = myToken, bucket = "mybucket")
resp$status_code
#> [1] 200
```

::: callout-warning
**A bucket must be empty before it can be deleted.** Delete all objects with `deleteObject()` first. Attempting to delete a non-empty bucket returns a `409 Conflict`:

```{r}
#| eval: false
resp$status_code
#> [1] 409
resp$content$reason
#> [1] "Bucket not empty"
```
:::
