---
title: "Design Automation"
---

Need to convert a DWG to PDF without spinning up an AutoCAD licence? Design Automation is AutoDesk's answer to cloud-based CAD scripting. You submit a work item, AutoDesk runs it on their infrastructure, and the output lands in a destination URL you specify.

This chapter focuses on the most common use case: DWG → PDF.

```{mermaid}
flowchart LR
  A[DWG file\nat source URL] --> B[makePdf\nsubmit work item]
  B --> C{checkPdf}
  C -- pending / inprogress --> C
  C -- success --> D[PDF file\nat destination URL]
```

## Convert a DWG to PDF

`makePdf()` submits the conversion job. The token needs the `code:all` scope. This is specific to Design Automation and different from the data scopes used elsewhere.

::: callout-warning
**Both `source` and `destination` must be publicly accessible URLs.** Private Google Drive links, localhost paths, and short-lived signed S3 URLs won't work. The AutoDesk cloud workers need to be able to reach them directly. Use a public host or a pre-signed URL with plenty of time left on it.
:::

```{r}
#| eval: false
resp <- getToken(
  id     = Sys.getenv("client_id"),
  secret = Sys.getenv("client_secret"),
  scope  = "code:all"
)
myToken <- resp$content$access_token
```

```{r}
#| eval: false
mySource      <- "http://download.autodesk.com/us/samplefiles/acad/visualization_-_aerial.dwg"
myDestination <- "https://drive.google.com/folderview?id=0BygncDVHf60mTDZVNDltLThLNmM&usp=sharing"

resp         <- makePdf(source = mySource, destination = myDestination, token = myToken)
myWorkItemId <- resp$content$id
resp$content
#> $id
#> [1] "f47ac10b-58cc-4372-a567-0e02b2c3d479"
#>
#> $status
#> [1] "pending"
#>
#> $stats$timeQueued
#> [1] "2026-04-23T10:15:30.412Z"
```

## Poll for Completion

Use the work item `id` to check status. A simple DWG typically takes **15–30 seconds**; complex drawings with many layers or external references take longer.

```{r}
#| eval: false
repeat {
  resp <- checkPdf(id = myWorkItemId, token = myToken)
  cat("Status:", resp$content$status, "\n")
  if (resp$content$status == "success") break
  Sys.sleep(5)
}
#> Status: pending
#> Status: inprogress
#> Status: success
```

The full timing breakdown when it's done:

```{r}
#| eval: false
resp$content
#> $id
#> [1] "f47ac10b-58cc-4372-a567-0e02b2c3d479"
#>
#> $status
#> [1] "success"
#>
#> $stats$timeQueued
#> [1] "2026-04-23T10:15:30.412Z"
#>
#> $stats$timeDownloadStarted
#> [1] "2026-04-23T10:15:31.804Z"
#>
#> $stats$timeInstructionsStarted
#> [1] "2026-04-23T10:15:34.217Z"
#>
#> $stats$timeFinished
#> [1] "2026-04-23T10:15:48.993Z"
```

Once `status` is `"success"`, the PDF is waiting at your `destination` URL.

## Batch Conversion

Converting a folder of DWGs is just a loop:

```{r}
#| eval: false
dwg_urls  <- c(
  "https://files.example.com/floor_plan.dwg",
  "https://files.example.com/site_plan.dwg",
  "https://files.example.com/elevations.dwg"
)
dest_base <- "https://output.example.com/pdfs/"

job_ids <- vapply(dwg_urls, function(url) {
  fname <- sub(".*/", "", url)
  resp  <- makePdf(source      = url,
                   destination = paste0(dest_base, sub("\\.dwg$", ".pdf", fname)),
                   token       = myToken)
  resp$content$id
}, character(1))

# Poll until all jobs complete
repeat {
  statuses <- vapply(job_ids, function(id) {
    checkPdf(id = id, token = myToken)$content$status
  }, character(1))
  cat(format(Sys.time(), "%H:%M:%S"), "—", paste(statuses, collapse = " | "), "\n")
  if (all(statuses == "success")) break
  Sys.sleep(10)
}
#> 10:20:05 — pending | pending | inprogress
#> 10:20:15 — success | inprogress | inprogress
#> 10:20:25 — success | success | success
```
