---
title: "Design Automation"
---

# Design Automation

The Design Automation API provides the ability to perform automated tasks on design files in the cloud.

## Convert a DWG File to a PDF File

Use `makePdf()` to convert a DWG file to a PDF file. The token must be generated with the `code:all` scope.

::: callout-warning
**Both `source` and `destination` must be publicly accessible URLs.** Private Google Drive links, localhost paths, or signed S3 URLs with expired signatures will cause the job to fail with a `400 Bad Request`. Use a public file host or a pre-signed URL with sufficient expiry time.
:::

```{r}
#| eval: false
resp <- getToken(
  id     = Sys.getenv("client_id"),
  secret = Sys.getenv("client_secret"),
  scope  = "code:all"
)
myToken <- resp$content$access_token
#> [1] "eyJhbGciOiJSUzI1NiIsImtpZCI6IlU3c1BKNVVpOWNaVjlIR2FhX1pIeDhEd3VYSHcifQ..."
```

```{r}
#| eval: false
mySource      <- "http://download.autodesk.com/us/samplefiles/acad/visualization_-_aerial.dwg"
myDestination <- "https://drive.google.com/folderview?id=0BygncDVHf60mTDZVNDltLThLNmM&usp=sharing"

resp <- makePdf(source = mySource, destination = myDestination, token = myToken)
myWorkItemId <- resp$content$id
resp$content
#> $id
#> [1] "f47ac10b-58cc-4372-a567-0e02b2c3d479"
#>
#> $status
#> [1] "pending"
#>
#> $stats
#> $stats$timeQueued
#> [1] "2026-04-23T10:15:30.412Z"
```

Check the status of the conversion using the WorkItem `id` returned by `makePdf()`:

```{r}
#| eval: false
resp <- checkPdf(id = myWorkItemId, token = myToken)
resp$content
#> $id
#> [1] "f47ac10b-58cc-4372-a567-0e02b2c3d479"
#>
#> $status
#> [1] "success"
#>
#> $stats
#> $stats$timeQueued
#> [1] "2026-04-23T10:15:30.412Z"
#>
#> $stats$timeDownloadStarted
#> [1] "2026-04-23T10:15:31.804Z"
#>
#> $stats$timeInstructionsStarted
#> [1] "2026-04-23T10:15:34.217Z"
#>
#> $stats$timeFinished
#> [1] "2026-04-23T10:15:48.993Z"
```

The response includes a `status` field (`pending`, `inprogress`, or `success`) and `stats` with timing information. Poll `checkPdf()` until `status` is `"success"`. Use the `aps_retry()` helper from the [Authentication](authentication.qmd#rate-limits) chapter to add automatic backoff:

```{r}
#| eval: false
repeat {
  resp <- checkPdf(id = myWorkItemId, token = myToken)
  cat("Status:", resp$content$status, "\n")
  if (resp$content$status == "success") break
  Sys.sleep(5)
}
#> Status: pending
#> Status: inprogress
#> Status: success
```

A typical conversion of a simple DWG takes **15–30 seconds**. Complex drawings with many layers or external references may take longer. The PDF will be written to the `destination` URL once the job completes.
