---
title: "Automated Drawing Reports"
---

You've got the layer summary, the attribute data, the charts. Now let's wrap it all into a polished, reproducible report that regenerates automatically whenever a new DWG lands in the bucket. This chapter combines `gt` [@R-gt] tables with Quarto's parameterised rendering to produce professional HTML reports with one function call.

## A gt Summary Table

`gt` turns a data frame into a presentation-ready table in a few chained calls. The `layer_summary` from [Layer Structure Analysis](layer-analysis.qmd) is a natural fit:

```{r}
#| eval: false
library(gt)

layer_summary |>
  gt() |>
  tab_header(
    title    = "Drawing Layer Summary",
    subtitle = "aerial.dwg — AutoDeskR extraction"
  ) |>
  cols_label(
    layer      = "Layer",
    n_objects  = "Objects",
    total_area = "Total Area",
    mean_area  = "Mean Area"
  ) |>
  fmt_number(columns = c(total_area, mean_area), decimals = 1) |>
  tab_style(
    style     = cell_fill(color = "#EBF3FB"),
    locations = cells_column_labels()
  ) |>
  tab_source_note(
    source_note = paste("Extracted", Sys.Date(), "via AutoDeskR v0.4.0")
  ) |>
  tab_options(table.font.size = "small")
```

## Parameterised Reports

A parameterised Quarto document accepts different inputs at render time, making it reusable for any DWG without editing the source file. Add a `params` block to the YAML:

``` yaml
---
title: "Drawing Report"
format: html
params:
  urn: "YOUR_URN_HERE"
  drawing_name: "aerial.dwg"
---
```

Access the parameters inside the document:

```{r}
#| eval: false
urn          <- params$urn
drawing_name <- params$drawing_name
```

Use `drawing_name` in headings so each rendered report is self-labelled:

```{r}
#| eval: false
cat("## Layer Summary:", drawing_name, "\n")
```

## Render Programmatically

`quarto::quarto_render()` renders the parameterised template to a named output file. Loop over your drawing list to generate a report for each one:

```{r}
#| eval: false
library(quarto)

drawings <- list(
  list(urn = urns[["site_plan.dwg"]],      name = "site_plan"),
  list(urn = urns[["floor_plan_v2.dwg"]], name = "floor_plan_v2")
)

for (dwg in drawings) {
  quarto_render(
    input          = "drawing-reports.qmd",
    output_file    = paste0("report_", dwg$name, ".html"),
    execute_params = list(
      urn          = dwg$urn,
      drawing_name = paste0(dwg$name, ".dwg")
    )
  )
  cat("Generated: report_", dwg$name, ".html\n", sep = "")
}
#> Generated: report_site_plan.html
#> Generated: report_floor_plan_v2.html
```

## Scheduling Reports

Keep reports fresh automatically by triggering `quarto_render()` on a schedule.

**GitHub Actions** — run on a cron or whenever DWGs are pushed:

``` yaml
# .github/workflows/drawing-report.yml
on:
  schedule:
    - cron: "0 6 * * 1"   # every Monday at 06:00 UTC
  push:
    paths: ["dwg-files/**"]

jobs:
  render:
    runs-on: ubuntu-latest
    steps:
      - uses: actions/checkout@v4
      - uses: r-lib/actions/setup-r@v2
      - run: Rscript -e 'quarto::quarto_render("drawing-reports.qmd")'
```

**cronR (local)** — schedule from a local R session:

```{r}
#| eval: false
library(cronR)
cmd <- cron_rscript("render_reports.R")
cron_add(command     = cmd,
         frequency   = "weekly",
         at          = "06:00",
         id          = "drawing_report",
         description = "Weekly DWG report")
```

::: callout-tip
Store credentials (`client_id`, `client_secret`, bucket name) as GitHub Actions secrets or in `~/.Renviron` on the local machine. Never hardcode them in the report source file. Your future self will thank you.
:::
