---
title: "Live Dashboards"
---

Here's where everything comes together: the 3D viewer from the [Viewer](viewer.qmd) chapter on the left, live sensor charts from [Linking Sensor Streams](sensor-streams.qmd) on the right, all in one Shiny app that refreshes automatically. This is what a digital twin dashboard looks like in R.

```{mermaid}
graph LR
  A[getToken] --> B[viewer3D\nAutoDesk Viewer]
  A --> C[Tandem API\nhttr2]
  C --> D[dygraph\ntime-series chart]
  B --> E[Shiny Dashboard]
  D --> E
```

## Required Packages

```{r}
#| eval: false
install.packages(c("shiny", "dygraphs", "xts", "httr2"))
```

## Helper: Fetch Stream Readings

Define a reusable function that the Shiny server calls reactively:

```{r}
#| eval: false
library(AutoDeskR)
library(httr2)
library(dygraphs)
library(xts)
library(shiny)

get_stream_readings <- function(facility_id, stream_id, hours_back = 24) {
  token <- getToken(
    id     = Sys.getenv("client_id"),
    secret = Sys.getenv("client_secret"),
    scope  = "data:read"
  )$content$access_token

  end_time   <- Sys.time()
  start_time <- end_time - hours_back * 3600
  resp <- request(
    paste0("https://tandem.autodesk.com/api/v1/twins/",
           facility_id, "/streams/", stream_id)
  ) |>
    req_url_query(
      start = format(start_time, "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
      end   = format(end_time,   "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")
    ) |>
    req_auth_bearer_token(token) |>
    req_perform()
  readings <- resp_body_json(resp)
  data.frame(
    timestamp = as.POSIXct(vapply(readings$values, `[[`, character(1), "t"),
                            format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
    value     = vapply(readings$values, `[[`, numeric(1), "v"),
    stringsAsFactors = FALSE
  )
}
```

## App Startup

Fetch the token and stream list once, before the UI and server are defined:

```{r}
#| eval: false
myEncodedUrn <- jsonlite::base64_enc(Sys.getenv("urn"))
myToken      <- getToken(id     = Sys.getenv("client_id"),
                         secret = Sys.getenv("client_secret"),
                         scope  = "data:read")$content$access_token
facilityId   <- "urn:adsk.dtdm:facility.abc123def456"

streams_resp <- request(
  paste0("https://tandem.autodesk.com/api/v1/twins/", facilityId, "/streams")
) |>
  req_auth_bearer_token(myToken) |>
  req_perform()
streams <- resp_body_json(streams_resp)

stream_choices <- setNames(
  vapply(streams, `[[`, character(1), "id"),
  vapply(streams, `[[`, character(1), "displayName")
)
```

## UI Layout

Viewer on the left, controls and chart on the right:

```{r}
#| eval: false
ui <- fluidPage(
  titlePanel("Digital Twin Dashboard"),
  fluidRow(
    column(7,
      viewer3D(urn        = myEncodedUrn,
               token      = myToken,
               viewerType = "headless")
    ),
    column(5,
      selectInput("stream_id", "Sensor stream:",
                  choices = stream_choices),
      sliderInput("hours_back", "Time window (hours):",
                  min = 1, max = 168, value = 24, step = 1),
      dygraphOutput("sensor_plot", height = "320px"),
      tableOutput("stats_table")
    )
  )
)
```

## Server Logic

A `reactiveTimer` polls Tandem every 60 seconds so the chart stays current without the user having to refresh the page:

```{r}
#| eval: false
server <- function(input, output, session) {

  auto_refresh <- reactiveTimer(60000)   # 60 seconds

  readings <- reactive({
    auto_refresh()   # re-run on the timer tick
    get_stream_readings(facility_id = facilityId,
                        stream_id   = input$stream_id,
                        hours_back  = input$hours_back)
  })

  output$sensor_plot <- renderDygraph({
    df   <- readings()
    ts   <- xts(df$value, order.by = df$timestamp)
    name <- names(stream_choices)[stream_choices == input$stream_id]
    dygraph(ts, main = name) |>
      dyRangeSelector() |>
      dyOptions(fillGraph = TRUE, fillAlpha = 0.15,
                drawGrid  = TRUE, colors   = "#367ABF") |>
      dyAxis("y", label = "Value")
  })

  output$stats_table <- renderTable({
    df <- readings()
    data.frame(
      Metric = c("Readings", "Mean", "Min", "Max"),
      Value  = c(nrow(df),
                 round(mean(df$value), 2),
                 round(min(df$value),  2),
                 round(max(df$value),  2))
    )
  }, striped = TRUE, bordered = TRUE, align = "lr")
}

shinyApp(ui, server)
```

::: callout-warning
The Shiny server fetches Tandem data from R, not from the user's browser, so CORS restrictions don't apply. The server process does need outbound HTTPS access to `tandem.autodesk.com`, so check this if you're deploying on a restricted corporate network.
:::

## Deploying to shinyapps.io

Keep credentials out of the app source. Pass them as server environment variables at deploy time:

```{r}
#| eval: false
library(rsconnect)

deployApp(
  appDir     = ".",
  appName    = "twin-dashboard",
  appEnvVars = list(
    client_id     = Sys.getenv("client_id"),
    client_secret = Sys.getenv("client_secret"),
    urn           = Sys.getenv("urn")
  )
)
```

