---
title: "MCP Server"
---

AI assistants are increasingly useful not just for answering questions but for *doing things*, calling APIs, reading files, running analyses. The Model Context Protocol (MCP) is the open standard that makes that possible. It defines how an AI assistant discovers and calls external tools during a conversation, in the same way a web browser discovers and calls REST APIs.

AutoDeskR functions are exactly the kind of tools an AI agent would want to call. "Summarise the layers in this DWG." "Translate this file and give me the bounding box." "Fetch the last 24 hours of HVAC sensor readings." Each of those is a handful of AutoDeskR function calls, and with an MCP server in front of them, an AI assistant can make those calls itself, without you writing a single line of R.

This chapter shows how to expose AutoDeskR as an MCP server so that Claude Desktop, VS Code Copilot, or any other MCP-compatible client can query BIM models directly.

## Proposed Tools

The table below maps ten useful MCP tools to their underlying AutoDeskR functions. Each tool is a thin wrapper. It handles authentication, calls the relevant function(s), and returns a structured JSON result.

| MCP Tool | Wraps | Description |
|------------------------|------------------------|------------------------|
| `get_token` | `getToken()` | Authenticate and return a bearer token |
| `upload_file` | `makeBucket()` + `uploadFile()` | Create a bucket and upload a local file |
| `translate_file` | `translateObj()` / `translateSvf()` + `checkFile()` | Translate a file and poll until complete |
| `get_layer_summary` | `translateSvf()` + `getData()` | Layer-level object count and area for a DWG |
| `get_object_tree` | `getObjectTree()` | Model hierarchy as a JSON tree |
| `get_metadata` | `getMetadata()` | List available metadata GUIDs for a translated model |
| `download_mesh` | `getOutputUrn()` + `downloadFile()` | Download the translated OBJ/STL to a local path |
| `make_pdf` | `makePdf()` + `checkPdf()` | Convert a DWG to PDF via Design Automation |
| `get_sensor_streams` | Tandem REST API | List sensor streams for a Tandem facility |
| `get_stream_readings` | Tandem REST API | Fetch time-series readings for a stream |

: {tbl-colwidths="[22,33,45]"}

## Setting Up a Local MCP Server

The [`mcptools`](https://posit-dev.github.io/mcptools/) [@R-mcptools] package turns R functions into MCP tools. Install it from CRAN:

```{r}
#| eval: false
install.packages("mcptools")
# development version: pak::pak("posit-dev/mcptools")
```

`mcptools` builds on [`ellmer`](https://ellmer.tidyverse.org/): each tool is an `ellmer::tool()` that wraps one or more AutoDeskR calls. Collect the tools in a single `.R` file that **returns a list** of them — `mcptools::mcp_server()` sources that file when the client starts the server. Because the file runs in a fresh R session, namespace everything (`ellmer::tool()`, not `tool()`).

Here's a working tools file with the `get_layer_summary` tool:

```{r}
#| eval: false
# autodeskr-mcp-tools.R
library(AutoDeskR)

# Authenticate once when the server starts, using environment variables
.token <- local({
  resp <- getToken(
    id     = Sys.getenv("client_id"),
    secret = Sys.getenv("client_secret"),
    scope  = "data:read data:write bucket:create"
  )
  resp$content$access_token
})

get_layer_summary <- ellmer::tool(
  function(urn) {
    encoded <- jsonlite::base64_enc(urn)

    # Retrieve the viewable GUID before fetching properties
    meta    <- getMetadata(urn = encoded, token = .token)
    guid    <- meta$content$data$metadata[[1]]$guid

    resp    <- getData(guid = guid, urn = encoded, token = .token)
    objects <- resp$content$data$collection

    layer_counts <- table(sapply(objects, `[[`, "name"))
    data.frame(layer = names(layer_counts),
               n_objects = as.integer(layer_counts))
  },
  name        = "get_layer_summary",
  description = "Return object count and total area by layer for a translated DWG URN.",
  arguments   = list(
    urn = ellmer::type_string("The raw objectId URN returned by uploadFile().")
  )
)
```

Add more tools by defining more `ellmer::tool()` objects, then return them all in a list at the end of the file. Here are two more useful examples:

```{r}
#| eval: false
# Tool: translate a file and poll until complete
translate_file <- ellmer::tool(
  function(urn) {
    encoded <- jsonlite::base64_enc(urn)
    translateObj(urn = encoded, token = .token)
    repeat {
      resp <- checkFile(urn = encoded, token = .token)
      if (resp$content$status == "success") break
      Sys.sleep(5)
    }
    list(status = "success", urn = encoded)
  },
  name        = "translate_file",
  description = "Translate a file URN to OBJ format and poll until complete.",
  arguments   = list(
    urn = ellmer::type_string("The raw objectId URN returned by uploadFile().")
  )
)

# Tool: return the model hierarchy as a JSON-serialisable list
get_object_tree <- ellmer::tool(
  function(urn) {
    encoded <- jsonlite::base64_enc(urn)
    meta    <- getMetadata(urn = encoded, token = .token)
    guid    <- meta$content$data$metadata[[1]]$guid
    resp    <- getObjectTree(guid = guid, urn = encoded, token = .token)
    resp$content$data$objects[[1]]
  },
  name        = "get_object_tree",
  description = "Return the layer/object hierarchy for a translated DWG URN.",
  arguments   = list(
    urn = ellmer::type_string("The raw objectId URN returned by uploadFile().")
  )
)

# The file must end by returning a list of every tool to expose
list(get_layer_summary, translate_file, get_object_tree)
```

The client launches the server with `Rscript`, pointing `mcp_server()` at this tools file (see the next section). To try it interactively first, call it directly — it blocks, waiting for a client to connect:

```{r}
#| eval: false
mcptools::mcp_server(tools = "autodeskr-mcp-tools.R")
```

## Connecting to Claude Desktop

Claude Desktop reads its MCP server list from a JSON config file. On any platform, the easiest way to open it is **Settings → Developer → Edit Config**, which reveals `claude_desktop_config.json`. To edit it directly instead, the file lives at:

| Platform | Config file path |
|:---------|:-----------------------------------------------------------------|
| macOS    | `~/Library/Application Support/Claude/claude_desktop_config.json` |
| Windows  | `%APPDATA%\Claude\claude_desktop_config.json`                     |
| Linux    | `~/.config/Claude/claude_desktop_config.json`                    |

: {tbl-colwidths="[20,80]"}

Add an entry for the AutoDeskR server. The command is `Rscript`, and the arguments tell `mcptools` to start a server with your tools file:

``` json
{
  "mcpServers": {
    "autodeskr": {
      "command": "Rscript",
      "args": ["-e", "mcptools::mcp_server(tools = '/path/to/autodeskr-mcp-tools.R')"],
      "env": {
        "client_id": "your_aps_client_id",
        "client_secret": "your_aps_client_secret"
      }
    }
  }
}
```

::: callout-important
**Replace before running:**

-   `/path/to/autodeskr-mcp-tools.R` — the absolute path to your tools file.
-   `your_aps_client_id` / `your_aps_client_secret` — your APS app credentials.

On Windows you may also need to give the full path to `Rscript.exe` as the `command`.
:::

Restart Claude Desktop. The AutoDeskR tools will appear in the tool palette and Claude will call them automatically when relevant.

## Example Agent Conversations

Once connected, you can have conversations like:

> **You:** I've uploaded `aerial.dwg` to bucket `acmecorp-dev`. The objectId is `urn:adsk.objects:os.object:acmecorp-dev/aerial.dwg`. Summarise the layers.

> **Claude:** *(calls `get_layer_summary`)* The drawing has 8 layers. The largest by object count is `A-DIMS` with 62 objects, followed by `A-TEXT` with 47. The largest by area is `A-SITE` at 156,823 square units.

> **You:** Compare the layer sets in `floor_plan_v1.dwg` and `floor_plan_v2.dwg`.

> **Claude:** *(calls `get_layer_summary` twice, compares results)* Version 2 adds `A-FURN` (34 objects) and `A-FURN-NEW` (12 objects). `A-TEMP-SITE` was present in v1 but has been removed.

> **You:** Fetch the last 24 hours of temperature readings from the HVAC sensor on level 2.

> **Claude:** *(calls `get_stream_readings`)* The stream `strm:temp-hvac-l2` shows a mean of 22.1 °C over the last 24 hours, ranging from 21.4 °C to 23.0 °C. Peak temperature was at 14:30 UTC.

## Security Considerations

A few things to keep in mind before deploying:

-   **Credentials:** Pass `client_id` and `client_secret` via environment variables (as shown in the config above), never hardcoded in the script. The server reads them at startup via `Sys.getenv()`.
-   **Token expiry:** APS tokens expire after 3600 seconds. For long-running servers, refresh the token periodically. Add a `Sys.time()` check in the handler and re-call `getToken()` if more than 50 minutes have elapsed.
-   **Rate limits:** APS enforces per-app rate limits. If the agent calls tools rapidly (e.g. in a loop over many files), add `Sys.sleep(0.5)` between calls. HTTP 429 responses should trigger a backoff, not a crash.
-   **Scope:** Request only the scopes the server actually needs. A read-only analysis server only needs `data:read`; it doesn't need `data:write` or `bucket:create`.
