---
title: "Reading OBJ and STL Meshes"
---

Once you've translated a CAD file through the [Model Derivative](model-derivative.qmd) API, you've got an OBJ or STL file sitting on disk. Now the fun part: loading it into R.

`rgl` [@R-rgl] and `Rvcg` [@R-Rvcg] both read OBJ and STL files. `rgl` is great for interactive viewing; `Rvcg` is the one to reach for when you want to compute things. It returns a `tmesh3d` object that plays nicely with the whole `Rvcg` family of analysis functions.

```{mermaid}
flowchart LR
  A[CAD file\nin OSS bucket] --> B[translateObj / translateStl\nModel Derivative]
  B --> C[downloadFile\narial.obj]
  C --> D[vcgImport\ntmesh3d object]
  D --> E[Analyse\nmetrics / visualise / compare]
```

## Prerequisites

```{r}
#| eval: false
install.packages(c("rgl", "Rvcg"))
```

## Getting the OBJ File

The full upstream chain, translate, poll, get output URN, download, picks up where the [Model Derivative](model-derivative.qmd) chapter left off:

```{r}
#| eval: false
library(AutoDeskR)

resp     <- getToken(id     = Sys.getenv("client_id"),
                     secret = Sys.getenv("client_secret"),
                     scope  = "data:read data:write")
myToken  <- resp$content$access_token

myEncodedUrn <- jsonlite::base64_enc(Sys.getenv("urn"))

translateObj(urn = myEncodedUrn, token = myToken)

repeat {
  status <- checkFile(urn = myEncodedUrn, token = myToken)
  if (status$content$status == "success") break
  Sys.sleep(5)
}

outputUrn       <- getOutputUrn(urn = myEncodedUrn, token = myToken)
myOutputUrn     <- outputUrn$content$derivatives[[1]]$children[[1]]$urn
myEncodedOutput <- jsonlite::base64_enc(myOutputUrn)

downloadFile(urn        = myEncodedUrn,
             output_urn = myEncodedOutput,
             token      = myToken,
             destfile   = "aerial.obj")
```

## Reading with Rvcg (recommended for analysis)

`vcgImport()` parses the OBJ and returns a `tmesh3d`, a list with a vertex matrix (`vb`) and a triangle index matrix (`it`). Every analysis function in the next three chapters works on this object.

```{r}
#| eval: false
library(Rvcg)
mesh_vcg <- vcgImport("aerial.obj")
str(mesh_vcg)
#> List of 4
#>  $ vb     : num [1:4, 1:2847] ...   # homogeneous vertex coords (X/Y/Z/W)
#>  $ it     : int [1:3, 1:1832] ...   # triangle face indices
#>  $ normals: num [1:4, 1:2847] ...
#>  $ material: list()
```

## Reading with rgl (for interactive viewing)

`readOBJ()` returns a similar `mesh3d` object that you can render immediately:

```{r}
#| eval: false
library(rgl)
mesh_obj <- readOBJ("aerial.obj")
```

::: callout-tip
OBJ files from the Model Derivative API sometimes contain multiple named objects (one per CAD layer or body). `vcgImport()` merges them into a single mesh by default. Use `rgl::readOBJ(material = TRUE)` if you need to keep them separate for per-layer analysis.
:::

## Reading STL

Exactly the same call works for STL files. `vcgImport()` handles both formats transparently:

```{r}
#| eval: false
mesh_stl <- vcgImport("aerial.stl")
# Same tmesh3d structure — all subsequent chapters work identically on OBJ and STL
```

## Quick Inspection

A few fast checks to make sure the mesh loaded correctly:

```{r}
#| eval: false
cat("Vertices:", ncol(mesh_vcg$vb), "\n")
#> Vertices: 2847
cat("Faces:   ", ncol(mesh_vcg$it), "\n")
#> Faces:    1832

# Coordinate ranges — useful sanity check that units are what you expect
apply(t(mesh_vcg$vb[1:3, ]), 2, range)
#>           [,1]      [,2]  [,3]
#> [1,] -142.830  -98.470  0.000
#> [2,]  318.620  241.190 12.500
```

The `mesh_vcg` object lives on from here through the [Mesh Metrics](mesh-metrics.qmd), [3D Visualisation](mesh-visualisation.qmd), and [Mesh Comparison](mesh-comparison.qmd) chapters, so keep it in your environment.
