---
title: "Model Derivative"
---

A DWG file sitting in a bucket is just bytes. The Model Derivative API is what turns those bytes into something useful, OBJ and STL meshes you can analyse in R, SVF viewables you can embed in a browser, or structured metadata you can query with `dplyr`. This chapter covers the full translation workflow.

The pattern is always the same: submit a job, poll until it finishes, then pull the output.

```{mermaid}
flowchart LR
  A[Upload file\nuploadFile] --> B[Translate\ntranslateObj / translateSvf]
  B --> C{checkFile}
  C -- pending / inprogress --> C
  C -- success --> D[Get output URN\ngetOutputUrn]
  D --> E[Download\ndownloadFile]
```

## Translate to OBJ

First get a token with `data:read` and `data:write` scopes. Note that only certain source formats can produce OBJ output (DWG, IPT, IAM, IPN, F3D, FBX. See the [APS supported formats list](https://aps.autodesk.com/en/docs/model-derivative/v2/developers_guide/supported-translations/) for the full matrix).

```{r}
#| eval: false
resp <- getToken(
  id     = Sys.getenv("client_id"),
  secret = Sys.getenv("client_secret"),
  scope  = "data:read data:write"
)
myToken <- resp$content$access_token
```

APS requires the file's URN to be Base-64 encoded before it'll accept it — `jsonlite::base64_enc()` handles that:

```{r}
#| eval: false
myEncodedUrn <- jsonlite::base64_enc(myUrn)
myEncodedUrn
#> [1] "dXJuOmFkc2sub2JqZWN0czpvcy5vYmplY3Q6bXlidWNrZXQvYWVyaWFsLmR3Zw=="
```

Kick off the translation:

```{r}
#| eval: false
resp <- translateObj(urn = myEncodedUrn, token = myToken)
resp$content
#> $result
#> [1] "created"
#>
#> $urn
#> [1] "dXJuOmFkc2sub2JqZWN0czpvcy5vYmplY3Q6bXlidWNrZXQvYWVyaWFsLmR3Zw=="
#>
#> $acceptedJobs$output$formats[[1]]$type
#> [1] "obj"
```

::: callout-warning
**OBJ translation only accepts certain source formats.** Submitting an unsupported format returns a `415 Unsupported Media Type`:

```{r}
#| eval: false
resp$status_code
#> [1] 415
resp$content$diagnostic
#> [1] "File type not supported for OBJ translation."
```
:::

Poll `checkFile()` until `status` hits `"success"`. The job can take anywhere from a few seconds to a few minutes depending on file complexity:

```{r}
#| eval: false
resp <- checkFile(urn = myEncodedUrn, token = myToken)
resp$content
#> $status
#> [1] "success"
#>
#> $progress
#> [1] "complete"
#>
#> $region
#> [1] "US"
```

::: callout-note
Translation status moves through `"pending"` → `"inprogress"` → `"success"`. Don't try to download before you see `"success"` — there's nothing there yet. Use the `aps_retry()` helper from the [Authentication](authentication.qmd#rate-limits) chapter to handle rate limiting while you wait.
:::

Retrieve the output URN, then download:

```{r}
#| eval: false
resp <- getOutputUrn(urn = myUrn, token = Sys.getenv("token"))
myOutputUrn        <- resp$content$derivatives[[1]]$children[[1]]$urn
myEncodedOutputUrn <- jsonlite::base64_enc(myOutputUrn)

resp <- downloadFile(
  urn        = myEncodedUrn,
  output_urn = myEncodedOutputUrn,
  token      = myToken,
  destfile   = "aerial.obj"
)
resp$status_code
#> [1] 200
```

## Translate to STL

Same pattern as OBJ, different function:

```{r}
#| eval: false
resp <- translateStl(urn = myEncodedUrn, token = myToken)
resp$content$acceptedJobs$output$formats[[1]]$type
#> [1] "stl"

# Poll until done
resp <- checkFile(urn = myEncodedUrn, token = myToken)
resp$content$status
#> [1] "success"
```

## Extract Metadata from a File

SVF format unlocks the structured metadata inside a model. The object tree, property sets, layer information. This is the format the Viewer also uses.

```{r}
#| eval: false
resp <- translateSvf(urn = myEncodedUrn, token = myToken)
resp$content$result
#> [1] "created"

# Wait for it...
resp <- checkFile(urn = myEncodedUrn, token = myToken)
resp$content$status
#> [1] "success"
```

`getMetadata()` returns the viewable GUIDs. You'll need the GUID for every subsequent metadata call:

```{r}
#| eval: false
resp   <- getMetadata(urn = myEncodedUrn, token = myToken)
myGuid <- resp$content$data$metadata[[1]]$guid
resp$content$data
#> $type
#> [1] "metadata"
#>
#> $metadata[[1]]$name
#> [1] "aerial"
#>
#> $metadata[[1]]$guid
#> [1] "a7b3c9d2-4e1f-4a8b-b6c2-9d3e7f5a1b4c"
#>
#> $metadata[[1]]$isMasterView
#> [1] TRUE
```

`getObjectTree()` returns the model's full object hierarchy. For a DWG, the children of the root node are its layers:

```{r}
#| eval: false
resp <- getObjectTree(guid = myGuid, urn = myEncodedUrn, token = myToken)
resp$content$data$objects[[1]]
#> $objectid
#> [1] 1
#>
#> $name
#> [1] "aerial.dwg"
#>
#> $objects[[1]]$name
#> [1] "Layer: A-SITE"
#>
#> $objects[[2]]$name
#> [1] "Layer: A-BLDG"
#>
#> $objects[[3]]$name
#> [1] "Layer: A-ROAD"
```

`getData()` pulls all geometry properties and material data, bounding boxes, layer assignments, custom attributes. This is the raw material for the [Layer Structure Analysis](layer-analysis.qmd) chapter:

```{r}
#| eval: false
resp <- getData(guid = myGuid, urn = myEncodedUrn, token = myToken)
resp$content$data$collection[[1]]
#> $objectid
#> [1] 2
#>
#> $name
#> [1] "Layer: A-SITE"
#>
#> $properties$`Layer and Material`$Layer
#> [1] "A-SITE"
#>
#> $properties$Geometry$`Bounding Box Min X`
#> [1] -142.83
#>
#> $properties$Geometry$`Bounding Box Max X`
#> [1]  318.62
```

::: callout-warning
**`getData()` can be large.** Complex models return thousands of objects in a single response. For very large DWGs, consider filtering the object tree first to identify only the layers or elements you care about.
:::
