---
title: "Model Derivative"
---

# Model Derivative

The Model Derivative API enables users to translate their designs into different formats and extract model data.

## Translate a File into OBJ Format

First get an access token with the `data:read` and `data:write` scopes. Note that only certain file types can be translated into OBJ format.

```{r}
#| eval: false
resp <- getToken(
  id     = Sys.getenv("client_id"),
  secret = Sys.getenv("client_secret"),
  scope  = "data:read data:write"
)
myToken <- resp$content$access_token
#> [1] "eyJhbGciOiJSUzI1NiIsImtpZCI6IlU3c1BKNVVpOWNaVjlIR2FhX1pIeDhEd3VYSHcifQ..."
```

The platform requires the file's URN to be Base-64 encoded:

```{r}
#| eval: false
myEncodedUrn <- jsonlite::base64_enc(myUrn)
myEncodedUrn
#> [1] "dXJuOmFkc2sub2JqZWN0czpvcy5vYmplY3Q6bXlidWNrZXQvYWVyaWFsLmR3Zw=="
```

Translate the file into OBJ format:

```{r}
#| eval: false
resp <- translateObj(urn = myEncodedUrn, token = myToken)
resp$content
#> $result
#> [1] "created"
#>
#> $urn
#> [1] "dXJuOmFkc2sub2JqZWN0czpvcy5vYmplY3Q6bXlidWNrZXQvYWVyaWFsLmR3Zw=="
#>
#> $acceptedJobs
#> $acceptedJobs$output
#> $acceptedJobs$output$formats[[1]]$type
#> [1] "obj"
```

::: callout-warning
**OBJ translation supports a limited set of source formats.** Supported types include IAM, IPT, IPN, DWG, F3D, and FBX. Submitting an unsupported format (e.g. a raw PDF or image) returns a `415 Unsupported Media Type`:

```{r}
#| eval: false
resp$status_code
#> [1] 415
resp$content$diagnostic
#> [1] "File type not supported for OBJ translation."
```

See the [APS supported formats list](https://aps.autodesk.com/en/docs/model-derivative/v2/developers_guide/supported-translations/) for the full matrix.
:::

Check the status of the translation:

```{r}
#| eval: false
resp <- checkFile(urn = myEncodedUrn, token = myToken)
resp$content
#> $type
#> [1] "design"
#>
#> $hasThumbnail
#> [1] "true"
#>
#> $status
#> [1] "success"
#>
#> $progress
#> [1] "complete"
#>
#> $region
#> [1] "US"
#>
#> $urn
#> [1] "dXJuOmFkc2sub2JqZWN0czpvcy5vYmplY3Q6bXlidWNrZXQvYWVyaWFsLmR3Zw=="
```

::: callout-note
`status` progresses through `"pending"` → `"inprogress"` → `"success"`. Poll `checkFile()` until `status` is `"success"` before attempting to download. Use the `aps_retry()` helper from the [Authentication](authentication.qmd#rate-limits) chapter to handle rate limiting automatically.
:::

When translation is complete, retrieve the output URN of the translated file (different from the source URN). The response may include URNs for multiple output types (obj, svf, png) — find the one ending in `.obj`:

```{r}
#| eval: false
resp <- getOutputUrn(urn = myUrn, token = Sys.getenv("token"))
myOutputUrn <- resp$content$derivatives[[1]]$children[[1]]$urn
myOutputUrn
#> [1] "urn:adsk.viewing:fs.file:dXJuOmFkc2sub2JqZWN0czpvcy5vYmplY3Q6bXlidWNrZXQv
#>       YWVyaWFsLmR3Zw==/output/geometry/bc3339b2-73cd-4fba-9cb3-15363703a354.obj"
```

Download the OBJ file locally using the `destfile` parameter:

```{r}
#| eval: false
myEncodedOutputUrn <- jsonlite::base64_enc(myOutputUrn)
resp <- downloadFile(
  urn        = myEncodedUrn,
  output_urn = myEncodedOutputUrn,
  token      = myToken,
  destfile   = "aerial.obj"
)
resp$status_code
#> [1] 200
```

## Translate a File into STL Format

Use `translateStl()`, which follows the same pattern as `translateObj()`:

```{r}
#| eval: false
resp <- translateStl(urn = myEncodedUrn, token = myToken)
resp$content
#> $result
#> [1] "created"
#>
#> $urn
#> [1] "dXJuOmFkc2sub2JqZWN0czpvcy5vYmplY3Q6bXlidWNrZXQvYWVyaWFsLmR3Zw=="
#>
#> $acceptedJobs
#> $acceptedJobs$output$formats[[1]]$type
#> [1] "stl"
```

Check the status:

```{r}
#| eval: false
resp <- checkFile(urn = myEncodedUrn, token = myToken)
resp$content
#> $status
#> [1] "success"
#>
#> $progress
#> [1] "complete"
#>
#> $derivatives[[1]]$outputType
#> [1] "stl"
#>
#> $derivatives[[1]]$status
#> [1] "success"
```

## Extract Data from a File

After uploading a file and encoding its URN, translate it into SVF format to unlock metadata and geometry extraction:

```{r}
#| eval: false
resp <- translateSvf(urn = myEncodedUrn, token = myToken)
resp$content
#> $result
#> [1] "created"
#>
#> $urn
#> [1] "dXJuOmFkc2sub2JqZWN0czpvcy5vYmplY3Q6bXlidWNrZXQvYWVyaWFsLmR3Zw=="
#>
#> $acceptedJobs$output$formats[[1]]$type
#> [1] "svf"

resp <- checkFile(urn = myEncodedUrn, token = myToken)
resp$content$status
#> [1] "success"
```

Retrieve file metadata using `getMetadata()`, which returns the `type`, `name`, and `guid` of each viewable. Note the `guid` and store it in `.Renviron` for reuse:

```{r}
#| eval: false
resp   <- getMetadata(urn = myEncodedUrn, token = myToken)
myGuid <- resp$content$data$metadata[[1]]$guid
resp$content$data
#> $type
#> [1] "metadata"
#>
#> $metadata
#> $metadata[[1]]$name
#> [1] "aerial"
#>
#> $metadata[[1]]$role
#> [1] "3d"
#>
#> $metadata[[1]]$guid
#> [1] "a7b3c9d2-4e1f-4a8b-b6c2-9d3e7f5a1b4c"
#>
#> $metadata[[1]]$isMasterView
#> [1] TRUE
```

Retrieve the full object tree of the model:

```{r}
#| eval: false
resp <- getObjectTree(guid = myGuid, urn = myEncodedUrn, token = myToken)
resp$content$data$objects[[1]]
#> $objectid
#> [1] 1
#>
#> $name
#> [1] "aerial.dwg"
#>
#> $objects
#> $objects[[1]]$objectid
#> [1] 2
#> $objects[[1]]$name
#> [1] "Layer: A-SITE"
#>
#> $objects[[2]]$objectid
#> [1] 3
#> $objects[[2]]$name
#> [1] "Layer: A-BLDG"
#>
#> $objects[[3]]$objectid
#> [1] 4
#> $objects[[3]]$name
#> [1] "Layer: A-ROAD"
```

Extract all geometry properties and material data from the model:

```{r}
#| eval: false
resp <- getData(guid = myGuid, urn = myEncodedUrn, token = myToken)
resp$content$data$collection[[1]]
#> $objectid
#> [1] 2
#>
#> $name
#> [1] "Layer: A-SITE"
#>
#> $properties
#> $properties$`Layer and Material`$Layer
#> [1] "A-SITE"
#>
#> $properties$`Layer and Material`$Material
#> [1] "ByLayer"
#>
#> $properties$Geometry$`Bounding Box Min X`
#> [1] -142.83
#>
#> $properties$Geometry$`Bounding Box Min Y`
#> [1]  -98.47
#>
#> $properties$Geometry$`Bounding Box Max X`
#> [1]  318.62
#>
#> $properties$Geometry$`Bounding Box Max Y`
#> [1]  241.19
```
