---
title: "Reality Capture"
---

Got a folder of overlapping photos and want a 3D mesh? That's the Reality Capture API (also called Photo-to-3D). You upload the images, AutoDesk's photogrammetry engine does the heavy lifting, and you download a georeferenced 3D model. No specialised hardware or software required.

Common uses:

- **As-built documentation** — fly a drone over a construction site each week and generate a mesh for each visit
- **Heritage recording** — photograph a structure with a phone and get a measurable 3D record
- **Site topography** — turn aerial photos into a terrain mesh for civil design
- **Mesh analysis** — feed the output into `Rvcg` via the [3D Geometry](mesh-reading.qmd) chapters

The full workflow goes like this:

```{mermaid}
flowchart LR
  A[Photos] --> B[createPhotoscene]
  B --> C[uploadImages]
  C --> D[processPhotoscene]
  D --> E{waitForPhotoscene\nor checkPhotoscene}
  E -- progress < 100 --> E
  E -- progress = 100 --> F[Download\nmesh / point cloud]
```

## Before You Start

::: callout-tip
**Good photos make good meshes.** A few guidelines that will save you a lot of frustration:

-   At least **20 images** per scene — 50–200 is typical for a small structure
-   **60–80 % overlap** between adjacent shots; every surface should appear in at least 3 photos
-   Consistent, diffuse lighting — harsh shadows and mirror-like reflections confuse the algorithm
-   Move the camera *around* the subject, don't just zoom in from one spot
-   GPS/EXIF tags are optional but improve georeferencing accuracy
:::

## Step 1: Authenticate

Reality Capture needs `data:read` and `data:write` scopes:

```{r}
#| eval: false
library(AutoDeskR)

resp    <- getToken(
  id     = Sys.getenv("client_id"),
  secret = Sys.getenv("client_secret"),
  scope  = "data:read data:write"
)
myToken <- resp$content$access_token
```

## Step 2: Create a Photoscene

A **photoscene** is the container for your images and output. The `format` parameter controls what comes out the other end:

| Format | Output |
|------------------------------------|------------------------------------|
| `"rcm"` | AutoDesk ReCap project (default) |
| `"rcs"` | ReCap point cloud — feeds into [Point Cloud Analysis](point-cloud-analysis.qmd) |
| `"obj"` | Wavefront OBJ mesh — feeds into [Reading OBJ and STL Meshes](mesh-reading.qmd) |
| `"ortho"` | Orthophoto + DSM GeoTIFF |
| `"report"` | PDF quality report |

```{r}
#| eval: false
ps             <- createPhotoscene(name = "site-survey-2026", format = "obj", token = myToken)
myPhotosceneId <- ps$content$photoscene$photosceneid
myPhotosceneId
#> [1] "K2RDAPFIN0OPKGU9B65Q2TJD"
```

## Step 3: Upload Images

Pass a character vector of local file paths. AutoDeskR bundles them into a single multipart POST. JPEG and PNG both work.

```{r}
#| eval: false
image_files <- list.files("photos/", pattern = "\\.jpg$", full.names = TRUE)

imgs <- uploadImages(
  photoscene_id = myPhotosceneId,
  files         = image_files,
  token         = myToken
)
imgs$content$Files$file[[1]]
#> $filename
#> [1] "site_001.jpg"
#>
#> $fileid
#> [1] "A1B2C3D4E5F6G7H8"
#>
#> $filesize
#> [1] 4718592
```

::: callout-warning
**Free-tier limits:** Maximum **1,000 images per photoscene** and **10 photoscenes per month**. Exceeding the quota returns a `403 Forbidden`. Upgrade to a paid plan for higher limits.
:::

## Step 4: Start Processing

One call kicks off the reconstruction:

```{r}
#| eval: false
proc <- processPhotoscene(photoscene_id = myPhotosceneId, token = myToken)
proc$content$photoscene$progressmsg
#> [1] "Queued"
```

## Step 5: Wait for It

### Option A — Block until done (recommended)

`waitForPhotoscene()` polls on a schedule and returns when the job finishes. Set `verbose = TRUE` to watch the progress tick up:

```{r}
#| eval: false
done <- waitForPhotoscene(
  photoscene_id = myPhotosceneId,
  token         = myToken,
  interval      = 60,    # check every 60 seconds
  timeout       = 3600,  # give up after an hour
  verbose       = TRUE
)
#> Reality Capture progress: 0 - Queued
#> Reality Capture progress: 10 - Uploading files
#> Reality Capture progress: 35 - Calibrating cameras
#> Reality Capture progress: 60 - Building point cloud
#> Reality Capture progress: 85 - Generating mesh
#> Reality Capture progress: 100 - Done
```

### Option B — Manual polling

If you want more control, call `checkPhotoscene()` yourself:

```{r}
#| eval: false
status <- checkPhotoscene(photoscene_id = myPhotosceneId, token = myToken)
status$content$photoscene$progress
#> [1] "75"
status$content$photoscene$progressmsg
#> [1] "Processing geometry"
```

::: callout-warning
**Processing time scales with image count.** A 50-image scene typically takes 5–15 minutes; 500 images may take over an hour. Set `timeout` generously. There's nothing worse than a premature timeout on an 80-minute job.
:::

## Step 6: Download the Output

When `progress` hits `"100"`, retrieve the download URL and save the mesh locally:

```{r}
#| eval: false
output_url <- done$content$photoscene$scenelink

download.file(
  url      = paste0(output_url, "/output.obj"),
  destfile = "site-survey-2026.obj",
  headers  = c(Authorization = paste("Bearer", myToken))
)
```

From here, the OBJ file is ready for [Reading OBJ and STL Meshes](mesh-reading.qmd) or further translation via the [Model Derivative](model-derivative.qmd) API.
