---
title: "Reality Capture"
---

# Reality Capture

The Reality Capture API (also called Photo-to-3D) converts a set of overlapping photographs into a 3D mesh or point cloud. You upload images — from a phone, DSLR, or drone — and AutoDesk's cloud photogrammetry engine stitches them into a georeferenced 3D model you can download or feed into the viewer.

Common use cases include:

-   **As-built documentation** — capture a construction site weekly with a drone and generate a mesh for each visit
-   **Heritage and inspection** — record a structure or asset in 3D from photos taken by field staff
-   **Site topography** — convert aerial photography into a terrain mesh for use in civil design
-   **Integration with other APIs** — translate the output mesh into OBJ or STL via the [Model Derivative](model-derivative.qmd) API for further analysis in R

## Before You Start

::: callout-tip
**Image guidelines for best results:**

-   Minimum **20 images** per scene; 50–200 is typical for a small structure
-   **60–80 % overlap** between adjacent images — every part of the subject should appear in at least 3 photos
-   Consistent, diffuse lighting; avoid harsh shadows or direct sun reflections
-   Keep the camera moving around the object, not zooming in from a fixed point
-   GPS/EXIF tags are optional but improve georeferencing accuracy
:::

## Step 1: Authenticate

Reality Capture requires the `data:read` and `data:write` scopes:

```{r}
#| eval: false
library(AutoDeskR)

resp    <- getToken(
  id     = Sys.getenv("client_id"),
  secret = Sys.getenv("client_secret"),
  scope  = "data:read data:write"
)
myToken <- resp$content$access_token
#> [1] "eyJhbGciOiJSUzI1NiIsImtpZCI6IlU3c1BKNVVpOWNaVjlIR2FhX1pIeDhEd3VYSHcifQ..."
```

## Step 2: Create a Photoscene

A **photoscene** is the container for your images and the resulting 3D output. The `format` parameter determines what file type is produced:

| Format | Output |
|---|---|
| `"rcm"` | AutoDesk ReCap project (default) |
| `"rcs"` | ReCap point cloud scan |
| `"obj"` | Wavefront OBJ mesh |
| `"ortho"` | Orthophoto + DSM GeoTIFF |
| `"report"` | PDF quality report |

```{r}
#| eval: false
ps             <- createPhotoscene(name = "site-survey-2026", format = "obj", token = myToken)
myPhotosceneId <- ps$content$photoscene$photosceneid
ps$content
#> $photoscene
#> $photoscene$photosceneid
#> [1] "K2RDAPFIN0OPKGU9B65Q2TJD"
#>
#> $photoscene$scenename
#> [1] "site-survey-2026"
#>
#> $photoscene$scenelink
#> [1] "https://developer.api.autodesk.com/photo-to-3d/v1/photoscene/K2RDAPFIN0OPKGU9B65Q2TJD"
```

## Step 3: Upload Images

Pass a character vector of local file paths. AutoDeskR transfers them in a single multipart POST. JPEG and PNG are supported.

```{r}
#| eval: false
image_files <- c(
  "photos/site_001.jpg",
  "photos/site_002.jpg",
  "photos/site_003.jpg"
  # ... up to 1000 images per scene
)

imgs <- uploadImages(
  photoscene_id = myPhotosceneId,
  files         = image_files,
  token         = myToken
)
imgs$content
#> $Files
#> $Files$file
#> $Files$file[[1]]
#> $Files$file[[1]]$filename
#> [1] "site_001.jpg"
#>
#> $Files$file[[1]]$fileid
#> [1] "A1B2C3D4E5F6G7H8"
#>
#> $Files$file[[1]]$filesize
#> [1] 4718592
#>
#> $Files$file[[2]]$filename
#> [1] "site_002.jpg"
#>
#> $Files$file[[2]]$fileid
#> [1] "B2C3D4E5F6G7H8I9"
#>
#> $Files$file[[2]]$filesize
#> [1] 5242880
```

::: callout-warning
**Free-tier limits:** The APS free tier allows a maximum of **1 000 images per photoscene** and **10 photoscenes per month**. Exceeding these limits returns a `403 Forbidden` response. Upgrade to a paid plan for higher quotas.
:::

## Step 4: Start Processing

Once all images are uploaded, trigger the reconstruction:

```{r}
#| eval: false
proc <- processPhotoscene(photoscene_id = myPhotosceneId, token = myToken)
proc$content
#> $photoscene
#> $photoscene$photosceneid
#> [1] "K2RDAPFIN0OPKGU9B65Q2TJD"
#>
#> $photoscene$progress
#> [1] "0"
#>
#> $photoscene$progressmsg
#> [1] "Queued"
```

## Step 5: Poll for Completion

### Option A — Manual polling

```{r}
#| eval: false
status <- checkPhotoscene(photoscene_id = myPhotosceneId, token = myToken)
status$content$photoscene
#> $photosceneid
#> [1] "K2RDAPFIN0OPKGU9B65Q2TJD"
#>
#> $progress
#> [1] "75"
#>
#> $progressmsg
#> [1] "Processing geometry"
```

`progress` is a string from `"0"` to `"100"`. When it reaches `"100"` the model is ready.

### Option B — Block until done (recommended)

`waitForPhotoscene()` polls every `interval` seconds and returns when `progress == "100"`:

```{r}
#| eval: false
done <- waitForPhotoscene(
  photoscene_id = myPhotosceneId,
  token         = myToken,
  interval      = 60,     # poll every 60 s
  timeout       = 3600,   # give up after 1 hour
  verbose       = TRUE
)
#> Reality Capture progress: 0 - Queued
#> Reality Capture progress: 10 - Uploading files
#> Reality Capture progress: 35 - Calibrating cameras
#> Reality Capture progress: 60 - Building point cloud
#> Reality Capture progress: 85 - Generating mesh
#> Reality Capture progress: 100 - Done
done$content$photoscene$progress
#> [1] "100"
```

::: callout-warning
**Processing time** depends on image count and scene complexity. A 50-image scene typically takes **5–15 minutes**; 500 images may take over an hour. Set `timeout` to at least twice your expected processing time to avoid premature errors.
:::

## Step 6: Download the Output Mesh

Once `progress == "100"`, retrieve the download link from the completed photoscene response:

```{r}
#| eval: false
# The output URL is embedded in the final checkPhotoscene response
output_url <- done$content$photoscene$scenelink
#> [1] "https://developer.api.autodesk.com/photo-to-3d/v1/photoscene/K2RDAPFIN0OPKGU9B65Q2TJD/properties"

# Download the OBJ mesh
download.file(
  url      = paste0(output_url, "/output.obj"),
  destfile = "site-survey-2026.obj",
  headers  = c(Authorization = paste("Bearer", myToken))
)
```

The downloaded OBJ file can be imported into R with the `rgl` or `Rvcg` packages for 3D visualisation and analysis, or translated further via the [Model Derivative](model-derivative.qmd) API.
