---
title: "Linking Sensor Streams"
---

Each element in a Tandem facility can have sensor streams attached to it — named time-series channels that report temperature, CO₂, occupancy, energy consumption, or whatever the building management system sends. Because streams are keyed to the same `objectId` that `getObjectTree()` and `getData()` use, joining live sensor data to BIM metadata is just a regular table join.

This chapter uses the facility set up in [AutoDesk Tandem Overview](tandem-overview.qmd).

## Listing Streams for a Facility

```{r}
#| eval: false
library(AutoDeskR)
library(httr2)
library(dplyr)

resp       <- getToken(id     = Sys.getenv("client_id"),
                       secret = Sys.getenv("client_secret"),
                       scope  = "data:read")
myToken    <- resp$content$access_token
facilityId <- "urn:adsk.dtdm:facility.abc123def456"

streams_resp <- request(
  paste0("https://tandem.autodesk.com/api/v1/twins/", facilityId, "/streams")
) |>
  req_auth_bearer_token(myToken) |>
  req_perform()

streams <- resp_body_json(streams_resp)

streams_df <- lapply(streams, function(s) {
  data.frame(stream_id    = s$id,
             display_name = s$displayName,
             element_id   = s$elementId,   # objectId of the attached model element
             unit         = s$unit,
             last_reading = s$lastValue,
             stringsAsFactors = FALSE)
}) |> do.call(what = rbind)

streams_df
#>            stream_id    display_name element_id  unit last_reading
#> 1  strm:temp-hvac-l2  HVAC Temp L2          42    °C         22.4
#> 2  strm:co2-office-3  CO₂ Office 3         108   ppm        623.0
#> 3  strm:occ-lobby     Lobby Occupancy        15  pers         14.0
#> 4  strm:energy-main   Main Energy          201   kWh       8412.5
```

## Fetching Time-Series Readings

Request a time window using ISO 8601 timestamps as query parameters:

```{r}
#| eval: false
streamId   <- "strm:temp-hvac-l2"
end_time   <- Sys.time()
start_time <- end_time - 86400   # last 24 hours

readings_resp <- request(
  paste0("https://tandem.autodesk.com/api/v1/twins/",
         facilityId, "/streams/", streamId)
) |>
  req_url_query(
    start = format(start_time, "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
    end   = format(end_time,   "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")
  ) |>
  req_auth_bearer_token(myToken) |>
  req_perform()

readings <- resp_body_json(readings_resp)

readings_df <- data.frame(
  timestamp = as.POSIXct(vapply(readings$values, `[[`, character(1), "t"),
                          format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
  value     = vapply(readings$values, `[[`, numeric(1), "v"),
  stringsAsFactors = FALSE
)

head(readings_df)
#>             timestamp value
#> 1 2026-04-23 06:00:00  21.8
#> 2 2026-04-23 06:15:00  22.1
#> 3 2026-04-23 06:30:00  22.4
```

## Plotting Readings

```{r}
#| echo: false
library(ggplot2)
set.seed(7)
readings_df <- data.frame(
  timestamp = seq(as.POSIXct("2026-04-23 06:00", tz = "UTC"),
                  by = "15 min", length.out = 96),
  value     = 21 + cumsum(rnorm(96, 0, 0.15))
)
```

```{r}
ggplot(readings_df, aes(x = timestamp, y = value)) +
  geom_line(colour = "#367ABF", linewidth = 0.8) +
  geom_ribbon(aes(ymin = value - 0.5, ymax = value + 0.5),
              fill = "#367ABF", alpha = 0.15) +
  labs(title    = "HVAC Temperature — Level 2",
       subtitle = "Last 24 hours",
       x = NULL, y = "Temperature (°C)") +
  theme_minimal()
```

## Joining Sensor Data to Model Metadata

`element_id` in the streams response equals `objectid` from `getData()`. A left join links live readings to layer, classification, and any custom attributes:

```{r}
#| eval: false
# props_df from the Layer Structure Analysis chapter

joined <- streams_df |>
  inner_join(props_df |> select(objectid, layer),
             by = c("element_id" = "objectid"))

joined
#>            stream_id    display_name element_id   unit layer
#> 1  strm:temp-hvac-l2  HVAC Temp L2          42    °C  M-HVAC
#> 3  strm:occ-lobby     Lobby Occupancy        15  pers A-LBBY
```

## Fetching All Streams at Once

Loop over the stream list to build a single tidy data frame of all readings:

```{r}
#| eval: false
all_readings <- lapply(streams_df$stream_id, function(sid) {
  resp <- request(
    paste0("https://tandem.autodesk.com/api/v1/twins/",
           facilityId, "/streams/", sid)
  ) |>
    req_url_query(
      start = format(start_time, "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
      end   = format(end_time,   "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")
    ) |>
    req_auth_bearer_token(myToken) |>
    req_perform()
  readings <- resp_body_json(resp)
  data.frame(stream_id = sid,
             timestamp = as.POSIXct(
               vapply(readings$values, `[[`, character(1), "t"),
               format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
             value     = vapply(readings$values, `[[`, numeric(1), "v"),
             stringsAsFactors = FALSE)
}) |> do.call(what = rbind)

# Join display names and units
all_readings <- all_readings |>
  left_join(streams_df |> select(stream_id, display_name, unit),
            by = "stream_id")
```

## Multi-Stream Overview Plot

```{r}
#| echo: false
all_readings <- data.frame(
  stream_id    = rep(c("strm:temp","strm:co2","strm:occ","strm:energy"), each = 96),
  timestamp    = rep(readings_df$timestamp, 4),
  value        = c(readings_df$value,
                   600 + cumsum(rnorm(96, 0, 5)),
                   pmax(10 + cumsum(rnorm(96, 0, 1)), 0),
                   8000 + seq(0, 400, length.out = 96)),
  display_name = rep(c("HVAC Temp L2","CO\u2082 Office 3","Lobby Occupancy","Main Energy"), each = 96),
  unit         = rep(c("\u00b0C","ppm","pers","kWh"), each = 96)
)
```

```{r}
ggplot(all_readings, aes(x = timestamp, y = value, colour = display_name)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ paste0(display_name, " (", unit, ")"),
             scales = "free_y", ncol = 2) +
  labs(title = "All Sensor Streams — Last 24 Hours",
       x = NULL, y = NULL) +
  theme_minimal() +
  theme(legend.position = "none")
```

The `all_readings` data frame feeds directly into the interactive Shiny dashboard in [Live Dashboards](live-dashboards.qmd).
