---
title: "AutoDesk Tandem Overview"
---

A BIM model is a snapshot of a building at design time. A **digital twin** takes that same model and connects it to the building's live operational data — temperature sensors, occupancy counters, energy meters — so you can see what's actually happening inside, not just what was drawn. AutoDesk Tandem is APS's platform for building and querying digital twins.

This chapter introduces Tandem concepts and shows how to talk to the Tandem REST API directly from R using `httr2` [@R-httr2] and a standard APS token from `getToken()`. See [@autodesk-tandem] for the full API reference.

The relationship between a BIM model and a Tandem twin looks like this:

```{mermaid}
graph LR
  A[BIM Model\nRVT / DWG file] --> B[Model Derivative API\ngetObjectTree / getData]
  A --> C[Tandem\nDigital Twin]
  C --> D[Sensor Streams\ntemperature / CO₂ / energy]
  B -- objectId --> C
  D -- time-series data --> E[R Analysis\nhttr2 + dplyr]
  B -- metadata --> E
```

::: callout-tip
Tandem requires an active **AutoDesk Tandem subscription** in addition to an APS app registration. Authentication uses the same OAuth Bearer token as the rest of this book — no extra credential setup needed.
:::

## Twin vs. BIM Model

| | BIM Model | Digital Twin |
|---|---|---|
| Primary data | Geometry + properties | Geometry + live sensor streams |
| Update frequency | Per design revision | Continuous / real-time |
| Primary use | Design & construction | Operations & maintenance |
| APS API | Model Derivative | Tandem REST |
| R entry point | `getObjectTree()`, `getData()` | `httr2::request()` |

## Authentication

The Tandem API accepts the same Bearer token as all other APS services:

```{r}
#| eval: false
library(AutoDeskR)
library(httr2)

resp    <- getToken(id     = Sys.getenv("client_id"),
                    secret = Sys.getenv("client_secret"),
                    scope  = "data:read")
myToken <- resp$content$access_token
```

## Listing Facilities

A *facility* in Tandem is the digital twin of a building or site — the top-level container for model data and sensor streams. List all facilities your token can access:

```{r}
#| eval: false
facilities_resp <- request("https://tandem.autodesk.com/api/v1/groups") |>
  req_auth_bearer_token(myToken) |>
  req_perform()

facilities <- resp_body_json(facilities_resp)

facilities_df <- lapply(facilities, function(f) {
  data.frame(id = f$id, name = f$name, stringsAsFactors = FALSE)
}) |> do.call(what = rbind)

facilities_df
#>                                     id               name
#> 1  urn:adsk.dtdm:facility.abc123def456  Office Block A
#> 2  urn:adsk.dtdm:facility.xyz789uvw012  Warehouse Site B
```

## Facility Details

```{r}
#| eval: false
facilityId <- "urn:adsk.dtdm:facility.abc123def456"

facility_resp <- request(
  paste0("https://tandem.autodesk.com/api/v1/twins/", facilityId)
) |>
  req_auth_bearer_token(myToken) |>
  req_perform()

facility <- resp_body_json(facility_resp)
cat("Name:        ", facility$displayName, "\n")
#> Name:         Office Block A
cat("Created:     ", facility$createTime, "\n")
#> Created:      2025-11-04T09:22:14Z
cat("Linked URNs:", length(facility$links), "\n")
#> Linked URNs:  2
```

`facility$links` contains the APS Model Derivative URNs of the BIM models embedded in this twin — the same URNs you'd pass to `getObjectTree()` or `getData()`, which is what makes the `objectId` the key that links the two worlds together.

## Looking Up an Element

Any `objectId` from `getData()` can be looked up in Tandem to find the corresponding physical element and its classification:

```{r}
#| eval: false
element_resp <- request(
  paste0("https://tandem.autodesk.com/api/v1/twins/",
         facilityId, "/elements/", 42L)   # objectId 42
) |>
  req_auth_bearer_token(myToken) |>
  req_perform()

element <- resp_body_json(element_resp)
cat("Name:          ", element$displayName, "\n")
#> Name:           Level 2 HVAC Unit
cat("Classification:", element$classification, "\n")
#> Classification: MEP:HVAC:Air Handling Unit
```

The [Linking Sensor Streams](sensor-streams.qmd) chapter shows how to pull time-series readings for elements like this one.
